Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-05-09 paper/main.tex
figure code + related work + color sync

@@

\subsection{Sheaves}
Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in keeping track of sections of data over opensets. Sheaves are a way of expressing that subsets, distributed data, and streaming data are local sections that can be glued together. Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}. Here the sheaf 
\begin{equation}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
denotes a sheaf from opensets on the base space $\dbase$ to sets of sections on the bundle $\dtotal$. As a contravariant functor it preserves morphism s

@@

\subsection{Maps Between Spaces}
\note{add couple of sentences introducing image functors}
We propose that there must exist a continuous \textcolor{functor}{map between topological spaces} \vindexc 
\begin{equation}
  \vindexc: \opensetgc \rightarrow \opensetc 
\end{equation}
such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$. Given \vindex\ we can construct two functors \vindexpull\ and \vindexpush\ which transport sheaves

\begin{equation}
  \label{eq:atct:morphisms:xi}
\begin{tikzcd}[row sep=huge]
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
  \arrow[rr, "\vindexpullc", maps to, color=functor] &  & 
  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}  \\
  \opensetc 
  \arrow[d, "{\vindexpushc\sheafc_{\gbasec,  \gtotalc}}", dashed, maps to, color=sheaf] 
  \arrow[u, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] &  & 
  \opensetgc 
  \arrow[d, "{\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] 
  \arrow[u, "{\vindexpullc\sheafc_{\dbasec, \dtotalc}}"', dashed, maps to, color=sheaf] 
  \arrow[ll, "\vindexc"', maps to, color=functor] \\
  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
  \arrow[ll, "\vindexpushc"', maps to, color=functor]                                                         
  \end{tikzcd}
\end{equation}
such that there is an association between graphic sections \gsection\ that take as input \gbasepoint\ and data sections that take input \dbasepointc\ when $\vindex(\gbasepoint) = \dbasepoint$. 
The \textcolor{functor}{pullback functor} $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
\begin{equation}
  \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
\end{equation}
such that there is a way to then look up what data values correspond with a graphic index
\begin{equation}
  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindex(\gbasepointc)) = \dsectionc(\dbasepointc)
\end{equation}

 The \textcolor{functor}{pushforward functor} $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
 \begin{equation}  
  \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
\end{equation}
such that it provides a way to look up which graphic  corresponds with a data index
\begin{equation}
  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \gsectionc(\gbasepointc)\;\forall \gbasepointc\in \vindexprec(\dbasepointc)
\end{equation}

@@

\begin{equation}
  \label{eq:artist:natural_transform:inclusions}
  \begin{tikzcd}
    \opensetc_1 & \cgamma{\opensetc_{1}}{\dtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*"', color=set] \arrow[rr, "\vartistc", color=artist] &  & \cgamma{\opensetc_{1}}{\vindexpushc\gtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*", color=set] \\
      &  &  &  \\
    \opensetc_{2} \arrow[uu, "\iota", hook, color=base] & \cgamma{\opensetc_{2}}{\dtotalc\restruction_{\opensetc_2}} \arrow[rr, "\vartistc", color=artist]                        &  & \cgamma{\opensetc_{2}}{\vindexpushc\gtotalc\restruction_{\opensetc_2}}                      
    \end{tikzcd}
\end{equation}
 The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace. Since the input and output of the artist is always sections over the same subspace, this is why the by construction the artist preserves the bookkeeping provided through the use of sheaves.

The output of an artist \vartist\ is a restricted subset of graphic sections
\begin{equation}
  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \} 
\end{equation} that are by definition only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}. We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving. For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.

\subsection{Operators}
\subsubsection{More Base Spaces: +}
\subsubsection{More Fibers: \times}

@@

We formalize this structure preservation as equivariance, which is that for every morphism on the data $(\dfunch_{\dtotal}, \dfunct_{\dtotal})$ there is an equivalent morphism on the graphic  $(\dfunch_{\gtotal}, \dfunct_{\gtotal})$ The artist is an equivariant map if the diagram commutes for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
\label{sec:artist:equivariance}
\begin{equation*}
  \begin{tikzcd}[ampersand replacement=\&, column sep=small]
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
  \arrow[rrr, "\vartistc", color=artist] 
  \arrow[d, "\dfuncpullc_{\dtotalc}"', color=action] 
  \& \& \& 
  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
  \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
  \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc\restriction_{\opensetc^{\prime}}} 
  \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \& 
  \opensetc 
   \& 
  \opensetgc 
  \arrow[l, "\vindexc"', color=functor] 
  \& 
  \imartist{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
  \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
  \& 
  \opensetc^{\prime} 
  \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
  \& 
  \opensetgc^{\prime} 
  \arrow[l, "\vindexc"', color=functor] 
  \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
  \& \\
  \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
  \arrow[rrr, "\vartistc^{\prime}", color=artist]  
  \& \& \& 
  \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
  \end{tikzcd}
\end{equation*}
such that starting at an arbitrary data point $\dsectionc(\dbasepointc)$ and transforming it into a different data point and then into a graphic 
\begin{equation*}
  \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
\end{equation*}
is equivalent to transforming the original data point into a graphic and then transforming the graphic into another graphic. The function $\dfunch_{\gtotal}$ induces a change in graphic generating function that matches the change in data.

@@

\begin{equation}
  \label{eq:artist:inout:diagram}
  \begin{tikzcd}[row sep=huge]
    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
    \arrow[rr, "\vartistc", color=artist] 
    \arrow[d, "\equivc"', color=monoid] &  & 
    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
    \arrow[d, "render"] 
    \arrow[lld, "\extractmc"', color=monoid, dashed] \\
    {\textcolor{set}{Hom}(\opensetc, \measurec)}  &  & visualization 
    \arrow[ll, "measure",]
    \end{tikzcd}
\end{equation}
so instead we introduce a measurement extraction function $\extractm = measure\circ render$ that decomposes the rendered output into measurable components and returns a map from the data index into the measurement space
\begin{equation}
  \label{eq:artist:actual}
  \extractmc:\gsectionc \mapsto (\opensetc \xrightarrow{\extractmc_{\gsectionc}} \measurec)
\end{equation}
such that $\extractmc_{\gsectionc}(\dbasepoint)$ returns the measurements $\measurec_{\dbasepointc}$ of the visual element generated for the data at that point $\gsectionc|_{\vindexprec(\dbasepointc)}$.

We also introduce a function \equivc\ that maps data to the measurement space directly 
\begin{equation}
\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)
\end{equation}
such that $\equivc_{\dsection}(\dbasepointc)$ is the expected set of measurements $\measure_{\dbasepoint}$. The pair of \textcolor{monoid}{testing functions} (\equivc, \extractmc) can be used to test that the expected encoding of the data matches the actual encoding
\begin{equation}
    \equivc(\dsectionc)(\dbasepointc) = \extractmc(\vartistc(\dsectionc))(\dbasepointc) = \extractmc(\gsectionc\circ\vindexprec)(\dbasepointc)=\measurec_{\dbasepointc}
\end{equation}

An artist is equivariant when changes to expected and actual are equivariant. As introduced in \autoref{eq:atct:morphism:base}, the base space transformation \dfunch\ is invariant because $\dsection\restriction_{\openset} = \dsection\restriction_{\dfunch(\openset^{prime})}$. This means that, for all points in the data $\dbasepoint \in \dbase$, the measurement should not change if only the base space is transformed 
\begin{equation}
  \equivc(\vartistc(\dsectionc))(\dfunch(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
\end{equation}
On the other hand, a change in sections \autoref{eq:atct:morphism:fiber} induces an equivalent change in measurements
\begin{equation}
  \equivc(\dfunctc(\dsectionc))(\dbasepoint) = \dfunctc_{\measure}(\extractmc(\vartistc(\dsectionc))(\dbasepoint))
\end{equation}
The change in measurements $\dfunctc_{\measure}$ is defined by the developer as the symmetry between data and graphic that the artist is expected to preserve. For example a data to color transformation could be defined with the constraint that it must respect permutation since the data is categorical or with the constraint that it respect scaling if the data is quantative.

@@

\subsection{}
The transformation $\vchannel$ is a map from data sections to graphic sections
\begin{equation}
\nu: \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}
\end{equation}
such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$. A pair of transformations $(\vchannel, \vchannel^{\prime})$ are equivariant if 
\begin{equation}
  \dfunct_{\vtotal} \vchannel(\dsection(\dbasepoint)) = \vchannel^{\prime}(\dfunct_{\dtotal}(\dsection(\dbasepoint)))\; \forall \dbasepoint \in \dbase
\end{equation}
and a transformation $\vchannel$ is equivariant when $\vchannel = \vchannel^{\prime}$

\subsection*{Fabricating \vchannelc's}
A consequence of $\pi(\dtotal) = \pi(\vchannel(\dtotal))$ is that $\vchannelc$ can be constructed as a pointwise transformation such that  
\begin{equation}
  \label{eq:constrution:nu}
  \vchannelc: \dfiberc_{\dbasepointc} \rightarrow \vfiberc_{\dbasepoint}
\end{equation}

This property means that a point in a single data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a corresponding point in a visual fiber $\velement \in \vfiber_{\dbasepointc}$. We use this property to construct compositions of $\vchannelc$ functions 
\begin{equation}
  \label{eq:construction:nu:fabrication}
  \begin{tikzcd}[col sep=Huge]
    \dfiberc_{\dbasepointc} 
    \arrow[rr, "\vchannelc", color=artist] 
    \arrow[rrrr, "\vchannelc^{\prime\prime}", dashed, bend right, color=artist] &  & 
    \vfiberc_{\dbasepointc}\coloneqq{\dfiberc_{\dbasepointc}^{\prime}} 
    \arrow[rr, "\vchannelc^{\prime}", color=artist] &  & 
    \vfiberc^{\prime}_{\dbasepointc}
  \end{tikzcd}
\end{equation}
such that any new $\vchannel$ can be constructed as a composition of two or more $\vchannel$s.

\subsubsection{Combining \vchannelc's} 
We can add visual parameters as

>>>>

\subsection{Topology}
The topological properties of visualization data are ..., as described by Wilkenson\cite{wilkinsonGrammarGraphics2005}.

\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{continuity.png}
\end{figure}
Vsisualization libraries designed around continuity,  
Visualization library algorithms assume

\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{k_different_types.png}
\end{figure}
Fiber bundles are a mathematical structure that can be use to encode the topological properties of visualization data, as proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}. Broadly, fiber bundles

\subsection{Equivariance}
\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{equiviz.png}
\end{figure}

@@

\subsection{Sheaves}
\label{sec:atct:sheaves}
Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in keeping track of sections of data over opensets. Sheaves are a way of expressing that subsets, distributed data, and streaming data are local sections that can be glued together. Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}. Here the sheaf 
\begin{equation}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
denotes a sheaf from opensets on the base space $\dbase$ to sets of sections on the bundle $\dtotal$. As a contravariant functor it preserves morphism s

@@

\subsection{Maps Between Spaces}
The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encodes the topology of the data topology. The graphic sections, defined in \autoref{eq:atct:fb_graphic_section}, are sections of sheaves over open sets $\opensetg \subseteq \gbase$ in a base space \gbase\ that encodes the topology of the graphic. We propose that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc 
\begin{equation}
  \vindexc: \opensetgc \rightarrow \opensetc 
\end{equation}
such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.

Given the map between spaces \vindex\, a property of sheaves is that there exist image functors\cite{ImageFunctorsSheaves2021} functors \vindexpull\ and \vindexpush. The \textcolor{functor}{pullback functor} $\vindexpullc$ sends sheaves over $\dbase$ to $\gbase$ and the \textcolor{functor}{pushforward functor} $\vindexpushc$  sends sheaves over $\gbase$ to $\dbase$
\begin{equation}
  \label{eq:atct:morphisms:xi}
\begin{tikzcd}[row sep=huge]
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
  \arrow[rr, "\vindexpullc", maps to, color=functor] &  & 
  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}  \\
  \opensetc 
  \arrow[d, "{\vindexpushc\sheafc_{\gbasec,  \gtotalc}}", dashed, maps to, color=sheaf] 
  \arrow[u, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] &  & 
  \opensetgc 
  \arrow[d, "{\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] 
  \arrow[u, "{\vindexpullc\sheafc_{\dbasec, \dtotalc}}"', dashed, maps to, color=sheaf] 
  \arrow[ll, "\vindexc"', maps to, color=functor] \\
  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
  \arrow[ll, "\vindexpushc"', maps to, color=functor]                                                         
  \end{tikzcd}
\end{equation}
such that there is an association between graphic sections \gsection\ that take as input \gbasepoint\ and data sections that take input \dbasepointc\ when $\vindex(\gbasepoint) = \dbasepoint$.

The pull back $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
\begin{equation}
  \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
\end{equation}
such that there is a way to then look up what data values correspond with a graphic index
\begin{equation}
  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindex(\gbasepointc)) = \dsectionc(\dbasepointc)
\end{equation}

 The pushforward $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
 \begin{equation}  
  \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
\end{equation}
such that it provides a way to look up which graphic  corresponds with a data index
\begin{equation}
  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \gsectionc(\gbasepointc)\;\forall \gbasepointc\in \vindexprec(\dbasepointc)
\end{equation}

@@

\begin{equation}
  \label{eq:artist:natural_transform:inclusions}
  \begin{tikzcd}
    \opensetc_1 & \cgamma{\opensetc_{1}}{\dtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*"', color=set] \arrow[rr, "\vartistc_{\opensetc_1}", color=artist] &  & \cgamma{\opensetc_{1}}{\vindexpushc\gtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*", color=set] \\
      &  &  &  \\
    \opensetc_{2} \arrow[uu, "\iota", hook, color=base] & \cgamma{\opensetc_{2}}{\dtotalc\restruction_{\opensetc_2}} \arrow[rr, "\vartistc_{\opensetc_2}", color=artist]                        &  & \cgamma{\opensetc_{2}}{\vindexpushc\gtotalc\restruction_{\opensetc_2}}                      
    \end{tikzcd}
\end{equation}
 The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace. Since the input and output of the artist is always sections over the same subspace, this is why the by construction the artist preserves the bookkeeping provided through the use of sheaves.

The output of an artist \vartist\ is a restricted subset of graphic sections
\begin{equation}
  \begin{split}
  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &\coloneqq\\ 
  &\{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \} 
  \end{split}
\end{equation} 
that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}. We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving. For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.

\subsection{Artist Operators}
\subsubsection{Addition: +}
\label{sec:artist:addition}
 $\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) + \vartist^{b}(\dsectionc^{b})$ &\\

 with input 
 \begin{equation}
    \begin{tikzcd}[column sep=small]
      \dtotalc^{a} \arrow[dd, "\pi"', color=total]  &  & \dtotalc^{b} \arrow[dd, "\pi"', color=total]                              \\
      {}    &  & {}                                                           \\
      \dbasec^{a} \arrow[uu, "\dsectionc^{a}"', dashed, bend right, color=section] &  & \dbasec^{b} \arrow[uu, "\dsectionc^{b}"', dashed, bend right, color=section]
      \end{tikzcd}
  \end{equation}
such that output 
\begin{equation}
  \label{eq:artist:plus:output}
  \gsectionc(\gbasepointc) = \begin{cases} \gsectionc^{a}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{a}) \\
    \gsectionc^{b}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{b})
  \end{cases}
\end{equation}

\subsubsection{Multiplication: $\times$}
$\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) \times \vartist^{b}(\dsectionc^{b})$
with input 
\begin{equation}
  \begin{tikzcd}
    \dtotalc^{a} \arrow[rdd, "\pi"', color=total] & & \dtotalc^{b} \arrow[ldd, "\pi"', color=total] \\
       & & \\
      & \dbasec \arrow[luu, "\dsectionc^{a}"', dashed, bend right, color=section] 
      \arrow[ruu, "\dsectionc^{b}"', dashed, bend right, color=section] &                                 
    \end{tikzcd}
  \end{equation}
  \\
 The output 
\begin{equation}
  \{\gsectionc^{a}(\gbasepointc), \gsectionc^{b}(\gbasepointc)\}, \gbasepointc \in \vindexprec(\dbasec)
\end{equation} 
which would mean s

which would require returning two points in the fiber $\gbase \rightarrow \dfiber \times \dfiber$. This function is adjoint to the addition function in \autoref{sec:artist:addition} 
\begin{equation}
  hom(\dbase, \dfiber\times \dfiber) \simeq hom(\dbase \plus \dbase, \dfiber)
\end{equation}
such that the $\times$ operator can be implemented as a special case of \autoref{eq:artist:plus:output} where $\gbase^a$ is a copy of $\gbase$ associated with $\gsection^a$ and $\gbase^b$ is a copy of $\gbase$ associated with $\gsection^{b}$.

@@

We formalize this structure preservation as equivariance, which is that for every morphism on the data $(\dfunch_{\dtotal}, \dfunct_{\dtotal})$ there is an equivalent morphism on the graphic  $(\dfunch_{\gtotal}, \dfunct_{\gtotal})$ The artist is an equivariant map if the diagram commutes for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
\label{sec:artist:equivariance}
\begin{equation}
  \autoref{eq:artist:equivariance}
  \begin{tikzcd}[ampersand replacement=\&, column sep=small]
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
  \arrow[rrr, "\vartistc", color=artist] 
  \arrow[d, "\dfuncpullc_{\dtotalc}"', color=action] 
  \& \& \& 
  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
  \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
  \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc\restriction_{\opensetc^{\prime}}} 
  \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \& 
  \opensetc 
   \& 
  \opensetgc 
  \arrow[l, "\vindexc"', color=functor] 
  \& 
  \imartist{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
  \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
  \& 
  \opensetc^{\prime} 
  \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
  \& 
  \opensetgc^{\prime} 
  \arrow[l, "\vindexc"', color=functor] 
  \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
  \& \\
  \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
  \arrow[rrr, "\vartistc^{\prime}", color=artist]  
  \& \& \& 
  \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
  \end{tikzcd}
\end{equation}
such that starting at an arbitrary data point $\dsectionc(\dbasepointc)$ and transforming it into a different data point and then into a graphic 
\begin{equation*}
  \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
\end{equation*}
is equivalent to transforming the original data point into a graphic and then transforming the graphic into another graphic. The function $\dfunch_{\gtotal}$ induces a change in graphic generating function that matches the change in data.

@@

\begin{equation}
  \label{eq:artist:inout:diagram}
  \begin{tikzcd}[row sep=huge]
    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
    \arrow[rr, "\vartistc", color=artist] 
    \arrow[d, "\equivc"', color=monoid] &  & 
    \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
    \arrow[d, "render"] 
    \arrow[lld, "\extractmc"', color=monoid, dashed] \\
    {\textcolor{set}{Hom}(\opensetc, \measurec)}  &  & visualization 
    \arrow[ll, "measure",]
    \end{tikzcd}
\end{equation}
so instead we introduce a measurement extraction function $\extractm = measure\circ render$ that decomposes the rendered output into measurable components and returns a map from the data index into the measurement space
\begin{equation}
  \label{eq:artist:actual}
  \extractmc:\gsectionc \mapsto (\opensetc \xrightarrow{\extractmc_{\gsectionc}} \measurec)
\end{equation}
such that $\extractmc_{\gsectionc}(\dbasepoint)$ returns a quantification of the measurable components $\measurec_{\dbasepointc}$ of the visual element generated for the data at that point $\gsectionc|_{\vindexprec(\dbasepointc)}$.

We also introduce a function \equivc\ that maps data to the measurement space directly 
\begin{equation}
\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)
\end{equation}
such that $\equivc_{\dsection}(\dbasepointc)$ is the expected set of measurements $\measure_{\dbasepoint}$. The pair of \textcolor{monoid}{verification functions} (\equivc, \extractmc) can be used to test that the expected encoding $\equivb_{\dsection}$ of the data matches the actual encoding $\extractm_{\gsection}$ 
\begin{equation}
  \label{eq:artist:verification}
    \equivc(\dsectionc)(\dbasepointc) = \extractmc(\vartistc(\dsectionc))(\dbasepointc) = \extractmc(\gsectionc\circ\vindexprec)(\dbasepointc)=\measurec_{\dbasepointc}
\end{equation}

An artist is equivariant when changes to expected and actual are equivariant. As introduced in \autoref{eq:atct:morphism:base}, the base space transformation \dfunch\ is invariant because $\dsection\restriction_{\openset} = \dsection\restriction_{\dfunch(\openset^{\prime})}$. This means that, for all points in the data $\dbasepoint \in \dbase$, the measurement should not change if only the base space is transformed 
\begin{equation}
  \equivc(\vartistc(\dsectionc))(\dfunch(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
\end{equation}
On the other hand, a change in sections \autoref{eq:atct:morphism:fiber} induces an equivalent change in measurements
\begin{equation}
  \equivc(\dfunctc(\dsectionc))(\dbasepoint) = \dfunctc_{\measure}(\extractmc(\vartistc(\dsectionc))(\dbasepoint))
\end{equation}
The change in measurements $\dfunctc_{\measure}$ is defined by the developer as the symmetry between data and graphic that the artist is expected to preserve. For example a data to color transformation could be defined with the constraint that it must respect permutation since the data is categorical or with the constraint that it respect scaling if the data is quantative.

@@

\subsection{Transforming Data into Measurable Components}
The transformation $\vchannel$ is a map from data sections to graphic sections
\begin{equation}
\nu: \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}
\end{equation}
such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$. As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$.

A consequence of $\pi(\dtotal) = \pi(\vchannel(\dtotal))$ is that $\vchannelc$ can be constructed as a pointwise transformation such that  
\begin{equation}
  \label{eq:constrution:nu}
  \vchannelc: \dfiberc_{\dbasepointc} \rightarrow \vfiberc_{\dbasepoint}
\end{equation}
which means that means that a point in a single data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a corresponding point in a visual fiber $\velement \in \vfiber_{\dbasepointc}$. Since $\vchannel$ can be constructed as a pointwise function property, we can construct $\vchannelc$ functions 
\begin{equation}
  \label{eq:construction:nu:fabrication}
  \begin{tikzcd}[col sep=Huge]
    \dfiberc_{\dbasepointc} 
    \arrow[rr, "\vchannelc", color=artist] 
    \arrow[rrrr, "\vchannelc^{\prime\prime}", dashed, bend right, color=artist] &  & 
    \vfiberc_{\dbasepointc}\coloneqq{\dfiberc_{\dbasepointc}^{\prime}} 
    \arrow[rr, "\vchannelc^{\prime}", color=artist] &  & 
    \vfiberc^{\prime}_{\dbasepointc}
  \end{tikzcd}
\end{equation}
such that any new $\vchannel$ can be constructed as a composition of two or more $\vchannel$s.

\subsubsection{Transformer Equivariance}

A pair of transformations $(\vchannel, \vchannel^{\prime})$ are equivariant if 
\begin{equation}
  \dfunct_{\vtotal} \vchannel(\dsection(\dbasepoint)) = \vchannel(\dfunct_{\dtotal}(\dsection(\dbasepoint)))
\end{equation}
for all $\dbasepoint \in \dbase$.  Since the visual sections and data sections are over the same space $\dfunc$

\begin{equation}
  \label{def:construction:nu:validate}
  \begin{tikzcd}[column sep=4em]
    {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{b}} 
    \arrow[d, "\pi_a"', color=fiber] 
    \arrow[r, "\vchannelc_{ab}", color=artist] 
    \arrow[rr, "\equivc_{ab}", bend left, color=monoid]  & 
    {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{b}} 
    \arrow[d, "\pi_a", color=fiber] & 
    \measurec_{\dbasepointc}^{ab} 
    \arrow[d, "\measurec\restriction_a", color=set] \\
    \dfiberc_{\dbasepointc}^a 
    \arrow[r, "\vchannelc_{a}", dashed, color=artist] & 
    \vfiberc_{\dbasepointc}^a 
    \arrow[r, "\simeq", dotted]  & 
    \measurec_{\dbasepointc}^a   \\
    {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{c}} 
    \arrow[u, "\pi_a", color=fiber] 
    \arrow[r, "\vchannelc_{ac}", color=artist] 
    \arrow[rr, "\equivc_{ac}", bend right, color=monoid] & 
    {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{c}} 
    \arrow[u, "\pi_a"', color=fiber] & 
    \measurec_{\dbasepointc}^{ac} 
    \arrow[u, "\measurec\restriction_a"', color=set]            
    \end{tikzcd}
\end{equation}

\subsection{Compositing Measurable Components into Output}
The compositing function $\vmarkc$ transforms the sections of visual elements $\vsectionc$ into sections of graphics $\gsectionc$.
\begin{equation}
  \vmarkc: \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
\end{equation}
